---
title: "141_FINAL_PROJECT_Group_Four"
author: "GroupFour"
date: "3/23/2018"
output:
  pdf_document: default
  html_document: default
---
# Reading the data
```{r}
library(readxl)
library(dplyr)
eye <-  read_excel("~/Desktop/2.24_renamed_data.xlsx")
# basic statistics of the data
dim(eye)
length(unique(eye$Name_unique))  #233 included 
summary(eye$Length_follow_up)
```

# Data exploration 
## Number of people in each group

# PRE_Operation
```{r}

#__________________________________________________________________________________
#Pre MRD Symmetry
PreMRDdiff1 <- eye %>%
  filter(abs(PreMRDdiff)<1)
length(unique(PreMRDdiff1$Name_unique))  #128

#Pre MRD Symmetry and Pre PTB Symmetry
PrePTB1 <- eye %>%
  filter(abs(PreMRDdiff) <1 ) %>%
  filter(abs(PrePTBdiff) < 3.5)
length(unique(PrePTB1$Name_unique))  #109

#Pre MRD Symmetry and Pre PTB Asymmetry
PrePTB2 <- eye %>%
  filter(abs(PreMRDdiff)<1) %>%
  filter(abs(PrePTBdiff) >= 3.5)
length(unique(PrePTB2$Name_unique))  #19


#__________________________________________________________________________________
#Pre MRD Asymmetry
PreMRDdiff2 <- eye %>%
  filter(abs(PreMRDdiff) >= 1)
length(unique(PreMRDdiff2$Name_unique))  #105

#Pre MRD Asymmetry and Pre PTB Symmetry
PrePTB3 <- eye %>%
  filter(abs(PreMRDdiff) >= 1) %>%
  filter(abs(PrePTBdiff) < 3.5)
length(unique(PrePTB3$Name_unique))  #84

#Pre MRD Asymmetry and Pre PTB Asymmetry
PrePTB4 <-  eye %>%
  filter(abs(PreMRDdiff) >= 1) %>%
  filter(abs(PrePTBdiff) >= 3.5 )
length(unique(PrePTB4$Name_unique))  #21
```




#POST_Operation
```{r}
#__________________________________________________________________________________
# PostMRD Symmetry
PostMRDdiff1 <- eye %>%
  filter(abs(PostMRDdiff) < 1 )
length(unique(PostMRDdiff1$Name_unique))  #194

#PostMRD Symmetry and PostPTB Symmetry
PostPTB5 <- eye %>%
  filter(abs(PostMRDdiff) < 1) %>%
  filter(abs(PostPTBdiff) < 3.5) 
length(unique(PostPTB5$Name_unique))   #170

#PostMRD Symmetry and PostPTB Asymmetry
PostPTB6 <- eye %>%
  filter(abs(PostMRDdiff) < 1) %>%
  filter(abs(PostPTBdiff) >= 3.5)
length(unique(PostPTB6$Name_unique))   #24


#__________________________________________________________________________________
# PostMRD Asymmetry
PostMRDdiff2  <- eye %>%
  filter(abs(PostMRDdiff)  >= 1)
length(unique(PostMRDdiff2$Name_unique))  #39

# PostMRD Asymmetry and PostPTB Symmetry
PostPTB7 <- eye %>%
  filter(abs(PostMRDdiff)  >= 1) %>%
  filter(abs(PostPTBdiff) < 3.5)
length(unique(PostPTB7$Name_unique))  #32

# PostMRD Asymmetry and PostPTB Asymmetry
PostPTB8 <- eye %>%
  filter(abs(PostMRDdiff) >= 1)  %>%
  filter(abs(PostPTBdiff) >= 3.5 )
length(unique(PostPTB8$Name_unique))   #7
```



# Recoding the variables: categorize
```{r}
#Symmetry:0
#Asymmetry:1

#Post PTB difference 
dim(eye)
eye$PostPTBdiff_c<-""
for (i in (1:nrow(eye))){
  if (eye$PostPTBdiff[i] > -3.5 & eye$PostPTBdiff[i] < 3.5){
    eye$PostPTBdiff_c[i] <- 0
  }
  else {
     eye$PostPTBdiff_c[i] <- 1
  }
}
table(eye$PostPTBdiff_c)



#Pre PTB difference 
eye$PrePTBdiff_c<-""
for (i in (1:nrow(eye))){
  if (eye$PrePTBdiff[i] > -3.5 & eye$PrePTBdiff[i] < 3.5){
    eye$PrePTBdiff_c[i] <- 0
  }
  else {
     eye$PrePTBdiff_c[i] <- 1
  }
}

table(eye$PrePTBdiff_c)




#Pre MRD
eye$PreMRDdiff_c<-""
for (i in (1:nrow(eye))){
  if (eye$PreMRDdiff[i] > -1 & eye$PreMRDdiff[i] < 1){
    eye$PreMRDdiff_c[i] <- 0
  }
  else {
     eye$PreMRDdiff_c[i] <- 1
  }
}
table(eye$PreMRDdiff_c)
table(eye$PreMRDdiff_c)


eye$PostMRDdiff_c<-""
for (i in (1:nrow(eye))){
  if (eye$PostMRDdiff[i] > -1 & eye$PostMRDdiff[i] < 1){
    eye$PostMRDdiff_c[i] <- 0
  }
  else {
     eye$PostMRDdiff_c[i] <- 1
  }
}
table(eye$PostMRDdiff_c);typeof(eye$PostMRDdiff_c)


table(eye$Lateral); as.factor(eye$Lateral)
```


# Making symmetry as baseline, changing factor level
```{r}
# change to factors:
#predict asymmetry: want to reject the Null
eye$PostMRDdiff_c <- factor(eye$PostMRDdiff_c,levels = c(0,1),labels = c("symmetry","asymmetry"))
eye$PreMRDdiff_c <- factor(eye$PreMRDdiff_c,levels = c(0,1),labels = c("symmetry","asymmetry"))
eye$PostPTBdiff_c <- factor(eye$PostPTBdiff_c,levels = c(0,1),labels = c("symmetry","asymmetry"))
eye$PrePTBdiff_c <- factor(eye$PrePTBdiff_c,levels = c(0,1),labels = c("symmetry","aymmetry"))
table(eye$PostPTBdiff_c)


left_indexes<-seq(2,nrow(eye),2)
right_indexes<-seq(1,nrow(eye),2)
left_eye <- eye[left_indexes,];  dim(left_eye)    #subset the left eye 
right_eye <- eye[right_indexes,];  dim(right_eye)  #subset the right eye 

```


# Four spineplots
```{r}
library(ggmosaic)
library(ggplot2)
##### Post  PTB symmetry vs. Pre PTB symmetry
ggplot(data=eye)+geom_mosaic(aes(weight=1, x=product(PostPTBdiff_c, PrePTBdiff_c), fill=factor(PrePTBdiff_c)), na.rm=TRUE)+theme_classic()+xlab("Pre PTB")+ylab("Post PTB")+
  scale_fill_discrete(name="Post PTB Symmetry", labels=c("Asymmetry", "Symmetry"))+ggtitle("Post  PTB symmetry vs. Pre PTB symmetry")

##### Post  MRD symmetry vs. Pre MRD symmetry
ggplot(data=eye)+geom_mosaic(aes(weight=1, x=product(PostMRDdiff_c, PreMRDdiff_c), fill=factor(PreMRDdiff_c)), na.rm=TRUE)+theme_classic()+xlab("Pre MRD")+
  ylab("Post MRD")+scale_fill_discrete(name="Post MRD Symmetry", labels=c("Asymmetry", "Symmetry"))+ggtitle("Post  MRD symmetry vs. Pre MRD symmetry")

##### Post  PTB symmetry vs. Pre MRD symmetry
ggplot(data=eye)+geom_mosaic(aes(weight=1, x=product(PostPTBdiff_c, PreMRDdiff_c), fill=factor(PreMRDdiff_c)), na.rm=TRUE)+theme_classic()+xlab("Pre MRD")+
  ylab("Post PTB")+scale_fill_discrete(name="Post PTB Symmetry", labels=c("Asymmetry", "Symmetry"))+ggtitle("Post  PTB symmetry vs. Pre MRD symmetry")

##### Post PTB symmetry vs. Post MRD symmetry
ggplot(data=eye)+geom_mosaic(aes(weight=1, x=product(PostMRDdiff_c, PostPTBdiff_c), fill=factor(PostPTBdiff_c)), na.rm=TRUE)+
  theme_classic()+xlab("Post MRD")+ylab("Post PTB")+scale_fill_discrete(name="Post MRD Symmetry", labels=c("Asymmetry", "Symmetry"))+ggtitle("Post PTB symmetry vs. Post MRD symmetry")
```

# Logistic regression
```{r}
#note: we are using "right-eye" data: (right-left)
#Use the unique rows
cor.test(eye$PreMRDdiff,eye$PostMRDdiff)

##### Full model 
logm1 <- glm(PostPTBdiff_c~Age+Lateral+Gender+PrePTBdiff_c+PreMRDdiff_c+PostMRDdiff_c+PreMRDdiff_c*PostMRDdiff_c,family="binomial",data = right_eye)
summary(logm1)

##### Odds ratio
round(exp(cbind(Estimate=coef(logm1),confint(logm1))),2)

##### Interation effect
interaction.plot(right_eye$PreMRDdiff_c,right_eye$PostMRDdiff_c,right_eye$PostMRDdiff,main= "Interaction effect for Pre MRD_diff and Post MRD_diff") # The interaction effect is not significant. 
library(effects)
plot(allEffects(logm1),ask=FALSE)


##### Reduced model
logm2 <- glm(PostPTBdiff_c~Age+PrePTBdiff_c+PreMRDdiff_c+PostMRDdiff_c,
         family="binomial",data = right_eye)
summary(logm2)
round(exp(cbind(Estimate=coef(logm2),confint(logm2))),2)

##### Deviance for logistic regression:
#logm1_Full
R2_devianceFull <- 1 - 134.51/182.45 
R2_devianceFull    #0.2627569

#logn2_Reduced
R2_devianceReduced <- 1 - 135.64/182.74
R2_devianceReduced  #0.2577432

##### Confusion Matrix:
library(nnet)
mlogm <- multinom(PostPTBdiff_c~Age+PrePTBdiff_c+PostMRDdiff_c+PreMRDdiff_c,data=right_eye)
summary(mlogm)
p <- predict(mlogm,right_eye)
tab <- table(p, right_eye$PostPTBdiff_c)
tab
(10+193)/233


##### Pearson goodness of fit of logistic regression:
#calculate person chi-square for full model
print(paste("Pearsons X^2=",round(sum(residuals(logm1,type="pearson")^2),3)))
#critical value 
qchisq(0.95,224)    #df=n-p-1=233-8-1=224
233.325 < 259.9144 
#less than critical value
#fail to reject the null, conclude the logistic model fits the data

#drop insignificant variables:
#calculate person chi-square for reduced model:
print(paste("Pearsons X^2=",round(sum(residuals(logm2,type="pearson")^2),3)))
#critical value 
qchisq(0.95,228)     #df=n-p-1=233-4-1=228
217.329 < 264.224  
#less than critical value
#fail to reject the null, conclude the logistic model fits the data
```

# Logistic regression with Cross Validation
```{r}
set.seed(141)
library(pROC)
library(ROCR)
dim(right_eye)

#split data into training[80%] and testing[20%]  
training <- right_eye[sample(233*0.8),]
testing <- right_eye[sample(233*0.2),]

##### training 
model <- glm(PostPTBdiff_c~Age+PrePTBdiff_c+PreMRDdiff_c+PostMRDdiff_c,
         family="binomial",data = training)
summary(model)
anova(model,test = "Chisq")
#odds ratio
round(exp(cbind(Estimate=coef(model),confint(model))),2)

library(pscl)
pR2(model)
##### testing
fitted.results <- predict(model,newdata = subset(testing,select=c(4,26,27,28)),type="response")
fitted.results <- ifelse(fitted.results > 0.5,1,0)
testing$PostPTBdiff_c <- as.factor(testing$PostPTBdiff_c)
levels(testing$PostPTBdiff_c) <- c(0,1)
table(fitted.results,testing$PostPTBdiff_c )
(36+3)/(36+3+2+5)  #Accuracy: 0.8478261


p <- predict(model,newdata = subset(testing,select=c(4,26,27,28)),type="response")
pr <- prediction(p,testing$PostPTBdiff_c)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")

##### ROC curve
plot(prf,lwd=3,colorize=TRUE,main="Receiver Operating Characteristic (ROC) curve")
lines(x=c(0,1),y=c(0,1),col="black",lwd=2)
polygon(x=c(0,1), y=c(0,0), col="gray")

prf@x.values
auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc  # 0.8092105
```



# Multiple Linear Regression Model:
```{r}
# Use the post-operation PTB difference as the response variable: (Test eye-brows)

##### check assumptions:
shapiro.test(right_eye$Age)
attach(right_eye)
round(cor(cbind(PostPTBdiff,Age,PrePTBdiff,PreMRDdiff_c,PostMRDdiff_c)),2)
#scatterplot:
pairs(~PostPTBdiff+Age+PrePTBdiff+PreMRDdiff_c+PostMRDdiff_c)

##### full model
m1 <- lm(PostPTBdiff~Age+PrePTBdiff+PreMRDdiff+PostMRDdiff+Lateral+Gender,data = right_eye)
summary(m1)
anova(m1)


##### reduced model
m2 <- lm(PostPTBdiff~Age+PrePTBdiff+PreMRDdiff+PostMRDdiff+Length_follow_up,data = right_eye)
summary(m2)
anova(m2)
library(car)
# vif(m2)
par(mfrow=c(2,2))
plot(m2)

##### Compare two models
library(stargazer)
library(car)
stargazer(logm1,logm2,m1,m2,type="text")
          # dep.dep.var.labels = c("PostPTBdiff_c (categorical)","PostPTBdiff(numerical)"),
          # out="modelsCompare.txt")
```

# Model Comparison
```{r}
#table to compare
library(stargazer)
library(car)
stargazer(logm1,logm2,m1,m2,type="text")
          # dep.dep.var.labels = c("PostPTBdiff_c (categorical)","PostPTBdiff(numerical)"),
          # out="modelsCompare.txt")
```

#Add one more logistic regression [changer as the y variable]
```{r}
##### create changer variable
#changer: success after the surgery (1,0)
changer <- abs(as.numeric(right_eye$PrePTBdiff_c) - as.numeric(right_eye$PostPTBdiff_c))
table(changer)
right_eye <- cbind(right_eye,changer)


##### logistic regression
logm3 <- glm(changer ~Age+PrePTBdiff_c+PreMRDdiff_c, family="binomial",data = right_eye)
summary(logm3)

##### odds ratio
round(exp(cbind(Estimate=coef(logm3),confint(logm3))),2)

#If the patients are brow asymmtery before the surgery, they are 16.54 times more likely to be brow symmtery after surgery(is a changer)
```


# Random Forest
```{r}
set.seed(1)
##### Split the training and testing (0.8 and 0.2)
training <- right_eye[sample(233,187),]
testing <- right_eye[-sample(233,187),]
training <- training[,c(1,7:9,25)]
testing <- testing[,c(1,7:9,25)]
library("randomForest")
##### Develop a predictive model
##### vimp plot
fit <- randomForest(as.factor(PostPTBdiff_c) ~ PreMRD1 + PrePTB + PostMRD,
                    data=training,
                    importance=TRUE,
                    ntree=2000)
Prediction <- predict(fit, testing) 
varImpPlot(fit, type=2)
importance(fit)
##### Accuracy rate and confusion matrix
table(Prediction,testing$PostPTBdiff_c)
```





